\section*{\centerline{Acknowledgments}}

	First and foremost, I want to thank my advisor, George Kollios, for his guidance and continued support throughout my doctorate journey, from the day he accepted me to his lab to this moment.
	I am honored to call him and his family my friends.
	I also thank Leo Reyzin for the time and effort he spent helping me with my work, and for the many cups of coffee and long life talks we have had over these years.
	I want to especially note the many hours-long philosophical discussions with Manos Athanassoulis --- the discussions I truly enjoyed.
	In the heat of our debates we have lived up to the Ph.\ part of our degrees.
	My special thanks go to Adam O'Neil for his countless contributions to my works.
	He has been an invaluable co-author, contributor, reviewer and friend.

	I would like to express my gratitude to my closest ally in this life, my beloved wife, Daria Bogatova.
	There is not a single piece of writing I did without her review, not a single figure I crafted without her keen eye refining the tiniest details, from fonts to colors.
	Even the source code of this very thesis has been branched off from her Bachelor work \cite{dasha-thesis}, and the names of my systems, \epsolute{} and \kanon{} are her inventions.
	She has always been the first to listen to my ideas, and the first to support them unconditionally.

	I would particularly like to thank my bosom friend, roommate and fellow Doctor, Oleksandr Narykov.
	Living in Boston will not be the same without him, and I hope his career will bring him here in no time.
	I also want to thank my good friend and fellow BU alumnus, Vasili Ramanishka.
	I miss our arguments and discussions in the middle of a night.

	My deepest appreciation goes to my colleagues and co-authors, Georgios Kellaris, Björn Tackmann, Kaoutar Elkhiyaoui, Angelo De Caro, Kobbi Nissim and Hamed Zamani.
	It has been a great pleasure and a rewarding experience working with you.
	Your input in my work has been indispensable.
	I am deeply grateful to my fellow Amazonians, Kiran Chinta, Sriram Krishnamurthy, Naresh Chainani, Ramchandra Kulkarni and Abhishek Rai Sharma with whom I shared a number of internships.
	I am very excited to join the team full-time!

	I owe my deepest gratitude to my family, my parents, grandma and a little sister.
	I thank you for always being there for me.
	For attending my defense and admiring my presentation even though the content may have been very technical and in a foreign language.
	I am also infinitely glad that my family has expanded during my doctorate program.
	I thank Dasha's side of the family, in the similar composition of parents, grandma and a little sister, for accepting me and supporting me through my journey.

	Last but not least, I want to note my little fluffy friend, Pixel.
	My years of the program would not have been the same without him sleeping on my keyboard.
