\chapter{Range queries in the snapshot model}\label{section:range-snapshot}
\thispagestyle{myheadings}

	\acrfull{ope} and \acrfull{ore} are two important encryption schemes that have been proposed in the area of query evaluation over encrypted data.
	These schemes can provide very efficient query execution but at the same time may leak some information to adversaries.
	In this chapter, we present the first comprehensive comparison among a number of important \acrshort{ope} and \acrshort{ore} schemes using a framework that we developed.
	We evaluate protocols that are based on these schemes as well.
	We analyze and compare them both theoretically and experimentally and measure their performance over database indexing and query evaluation techniques using not only execution time but also \acrshort{io} performance and usage of cryptographic primitive operations.
	Our comparison reveals some interesting insights concerning the relative security and performance of these approaches in database settings.
	Furthermore, we propose a number of improvements for some of these scheme and protocols.
	Finally, we provide a number of suggestions and recommendations that can be valuable to database researchers and users.

	\emph{Some of the following sections were paraphrased or taken verbatim from the following published work.}

	\cite{ore-benchmark-17} \printpublication{ore-benchmark-17}

	\input{sections/snapshot-range-queries/introduction}

	\input{sections/snapshot-range-queries/security}

	\input{sections/snapshot-range-queries/protocols/protocols}

	\input{sections/snapshot-range-queries/evaluation}

	\input{sections/snapshot-range-queries/conclusion}
